Meta-analysis fear learning

### 2022-03-01 manuscript/manuscript.Rmd
update 2022-03-01 17:37:58

To examine the effect of a categorical variable,
a reference category must be chosen.
Dummy variables encode the difference between each remaining category and this reference category.
When examining the results,
the intercept represents the expected effect size for a study that falls within the reference category for all categorical variables.
The effect of dummy variables represents the difference of that category with the reference category.
If a dummy variable has a significant effect, that means that that group's mean differs significantly from the reference category's mean (i.e., from the intercept).

```{r, eval = run_everything}
library(metacart)
modres <- lapply(processes, function(p){
  df <- dat[dat$Process == p, ]
  catvars <- sapply(df, class) == "factor" & names(df) %in% mods
  df[catvars] <- lapply(df[catvars], function(v){
    tab <- table(v)
    relevel(v, ref = names(tab)[which.max(tab)])
  })
  refcat <- data.frame(do.call(rbind,lapply(names(df[catvars]), function(n){c(n, levels(df[[n]])[1])})))
  names(refcat) <- c("Variable", "Reference")
  modmat <- model.matrix(~., df[mods])
  constcols <- colSums(modmat) == 0 | colSums(modmat) == nrow(modmat)
  constcolnames <- colnames(modmat)[constcols]
  modmat <- modmat[, !constcols]
  modmatlist <- as.list(data.frame(modmat))
  dupcols <- duplicated(modmatlist)
  modmat <- modmat[, !dupcols]
  these_dups <- modmatlist[dupcols]
  modmatlist <- as.list(data.frame(modmat))
  for(i in 1:length(these_dups)){
    idnum <- which(duplicated(c(modmatlist, these_dups[i]), fromLast = T))
    colnames(modmat)[idnum] <- paste0(colnames(modmat)[idnum], ";", names(these_dups)[i])
  }
  yi_notcentered <- df$yi# - mlm[[p]]$mod$b[1,1]
  res <- rma(yi = yi_notcentered,
             vi = df$vi,
             mods = modmat)
  vifs <- vif(res)
  saveRDS(res, paste0("rma_", p, ".RData"))
  tabres <- data.frame(Parameter = c(rownames(res$b), "Tau2"),
                       cbind(
                         Estimate = c(res$b, res$tau2),
                         se = c(res$se, res$se.tau2),
                         ci.lb = c(res$ci.lb, res$tau2-1.96*res$se.tau2),
                         ci.ub = c(res$ci.ub, res$tau2+1.96*res$se.tau2),
                         p = c(res$pval, res$QEp)))
  tabres$CI <- conf_int(lb = tabres$ci.lb, ub = tabres$ci.ub)
  tabres[c("ci.lb", "ci.ub")] <- NULL
  tabres$VIF <- c(NA, vifs$vif, NA)
  write.csv(tabres, paste0("rma_mods_", p, ".csv"))

@@

  pma <- brma(yi ~., study = "study", data = moddat, standardize = standardize)
  saveRDS(pma, paste0("pma_", p, ".RData"))
  sumpma <- data.frame(summary(pma)$coefficients)
  sumpma <- cbind(rownames(sumpma), sumpma)
  names(sumpma)[1:3] <- c("Parameter", "Estimate", "se")
  sumpma <- sumpma[!sumpma$Parameter == "tau", ]
  rownames(sumpma) <- NULL
  sumpma$CI <- conf_int(lb = sumpma$X2.5., ub = sumpma$X97.5.)
  pstars <- c("*", "")[as.integer(apply(sumpma[, c("X2.5.", 
                                                   "X97.5.")], 1, function(x) {
                                                     sum(sign(x)) == 0
                                                   })) + 1]
  sumpma$p <- pstars
  sumpma[c("sd", grep("^X", names(sumpma), value = T))] <- NULL
  write.csv(sumpma, paste0("brma_mods_", p, ".csv"))
  moddatvi <- moddat$vi
  moddat <- moddat[, -which(colnames(moddat) == "vi")]
  resmc <- suppressWarnings(metacart::FEmrt(reformulate(mods, response = "yi"), data = df, vi = df$vi))
  list(rma = tabres,
       brma = sumpma,
       mc = resmc,
       refcat = refcat)
})
names(modres) <- processes
saveRDS(modres, "modres.RData")
```

@@

<!-- predictie intervallen van elke categorie een goede manier is om de data te visualiseren en de data duidelijk over te brengen op de lezer. -->
```{r, eval = run_everything}
library(ggridges)
for(p in processes){
  res <- readRDS(paste0("pma_", p, ".RData"))
  pred <- pema:::predict.brma(res, newdata = res$X, type = "samples")
  for(v in c("ssri", "freq", "disease", "species", "test")){
    print(paste0(v, "_", p))
    catdf <- res$X[, startsWith(colnames(res$X), v), drop = F]
    splt <- apply(catdf, 1, which.max) + 1
    splt[rowSums(catdf == 0) == ncol(catdf)] <- 1
    labs <- c(modres[[p]]$refcat$Reference[modres[[p]]$refcat$Variable == v],
              unique(gsub(".", "", gsub(v, "", grep(paste0("^", v), colnames(res$X), value = T)), fixed = TRUE)))
    splt <- factor(splt, labels = labs)
    df_plot <- do.call(rbind, lapply(levels(splt), function(l){
      data.frame(Value = as.vector(pred[splt == l, ]),
                 Group = l)
    }))
    meds <- tapply(df_plot$Value, df_plot$Group, median)
    postplot <- ggplot(df_plot, aes(x = Value, y = Group)) +
      geom_density_ridges(scale = 1, alpha = 0) +
      geom_segment(data = data.frame(x = meds, xend = meds, y = 1:length(meds), yend = (2:(length(meds)+1))), aes(x = x, xend = xend, y = y, yend = yend), inherit.aes = F, colour = "skyblue3", size = 1) +
      theme_bw() +
      scale_y_discrete(expand =c(0,0))+
      labs(title = p,
           x = paste0("Posterior predicted effect size"),
           y = v)
    ggsave(filename = paste0("group_plot_", v, "_", p, ".png"), plot = postplot, device = "png")
  }
}
```

>>>>

To examine the effect of a categorical variable,
a reference category must be chosen.
Dummy variables encode the difference between each remaining category and this reference category.
When examining the results,
the intercept represents the expected effect size for a study that falls within the reference category for all categorical variables.
The effect of dummy variables represents the difference of that category with the reference category.
If a dummy variable has a significant effect, that means that that group's mean differs significantly from the reference category's mean (i.e., from the intercept).
The analyses below used the following reference categories:

```{r}
refcat <- structure(list(Variable = c("ssri", "freq", "disease", "species", "test"
), Reference = c("FLX", "chronic", "stress", "rat", "CF")), class = "data.frame", row.names = c(NA, -5L))

datatable(
  refcat,
  rownames = FALSE,
  caption = "Reference category for each variable")
```

```{r, eval = run_everything}
library(metacart)
modres <- lapply(processes, function(p){
  df <- dat[dat$Process == p, ]
  df[refcat$Variable] <- lapply(refcat$Variable, function(n){
    v <- df[[n]]
    v <- relevel(v, ref = refcat$Reference[refcat$Variable == n])
    droplevels(v)
  })
  # refcat <- data.frame(do.call(rbind,lapply(names(df[catvars]), function(n){c(n, levels(df[[n]])[1])})))
  # names(refcat) <- c("Variable", "Reference")
  df_mods <- df[mods]
  constcols <- sapply(df_mods, function(i){length(unique(i))}) == 1
  df_mods <- df_mods[, !constcols, drop = FALSE]
  modmat <- model.matrix(~., df_mods)
  constcols <- colSums(modmat) == 0 | colSums(modmat) == nrow(modmat)
  constcolnames <- colnames(modmat)[constcols]
  modmat <- modmat[, !constcols]
  modmatlist <- as.list(data.frame(modmat))
  dupcols <- duplicated(modmatlist)
  modmat <- modmat[, !dupcols]
  these_dups <- modmatlist[dupcols]
  modmatlist <- as.list(data.frame(modmat))
  for(i in 1:length(these_dups)){
    idnum <- which(duplicated(c(modmatlist, these_dups[i]), fromLast = T))
    colnames(modmat)[idnum] <- paste0(colnames(modmat)[idnum], ";", names(these_dups)[i])
  }
  yi_notcentered <- df$yi# - mlm[[p]]$mod$b[1,1]
  res <- rma(yi = yi_notcentered,
             vi = df$vi,
             mods = modmat)
  vifs <- vif(res)
  saveRDS(res, paste0("rma_", p, ".RData"))
  tabres <- data.frame(Parameter = c(rownames(res$b), "Tau2"),
                       cbind(
                         Estimate = c(res$b, res$tau2),
                         se = c(res$se, res$se.tau2),
                         ci.lb = c(res$ci.lb, res$tau2-1.96*res$se.tau2),
                         ci.ub = c(res$ci.ub, res$tau2+1.96*res$se.tau2),
                         p = c(res$pval, res$QEp)))
  tabres$CI <- conf_int(lb = tabres$ci.lb, ub = tabres$ci.ub)
  tabres[c("ci.lb", "ci.ub")] <- NULL
  tabres$VIF <- c(NA, vifs$vif, NA)
  write.csv(tabres, paste0("rma_mods_", p, ".csv"))

@@

  pma <- brma(yi ~., study = "study", data = moddat, standardize = standardize)
  saveRDS(pma, paste0("pma_", p, ".RData"))
  sumpma <- data.frame(summary(pma)$coefficients)
  sumpma <- cbind(rownames(sumpma), sumpma)
  names(sumpma)[1:3] <- c("Parameter", "Estimate", "se")
  sumpma <- sumpma[!sumpma$Parameter == "tau", ]
  rownames(sumpma) <- NULL
  sumpma$CI <- conf_int(lb = sumpma$X2.5., ub = sumpma$X97.5.)
  pstars <- c("*", "")[as.integer(apply(sumpma[, c("X2.5.", 
                                                   "X97.5.")], 1, function(x) {
                                                     sum(sign(x)) == 0
                                                   })) + 1]
  sumpma$p <- pstars
  sumpma[c("sd", grep("^X", names(sumpma), value = T))] <- NULL
  write.csv(sumpma, paste0("brma_mods_", p, ".csv"))
  moddatvi <- moddat$vi
  moddat <- moddat[, -which(colnames(moddat) == "vi")]
  resmc <- suppressWarnings(metacart::FEmrt(reformulate(mods, response = "yi"), data = df, vi = df$vi))
  list(rma = tabres,
       brma = sumpma,
       mc = resmc)
})
names(modres) <- processes
saveRDS(modres, "modres.RData")
```

@@

<!-- predictie intervallen van elke categorie een goede manier is om de data te visualiseren en de data duidelijk over te brengen op de lezer. -->
```{r, eval = run_everything}
refcat <- structure(list(Variable = c("ssri", "freq", "disease", "species", "test"
), Reference = c("FLX", "chronic", "stress", "rat", "CF")), class = "data.frame", row.names = c(NA, 
-5L))
library(ggridges)
for(p in processes){
  res <- readRDS(paste0("pma_", p, ".RData"))
  pred <- pema:::predict.brma(res, newdata = res$X, type = "samples")
  vars <- c("ssri", "freq", "disease", "species", "test")
  vars <- vars[sapply(vars, function(v){any(grepl(v, colnames(res$X)))})]
  for(v in vars){
    catdf <- res$X[, startsWith(colnames(res$X), v), drop = F]
    splt <- apply(catdf, 1, which.max) + 1
    splt[rowSums(catdf == 0) == ncol(catdf)] <- 1
    labs <- c(refcat$Reference[refcat$Variable == v],
              unique(gsub(".", "", gsub(v, "", grep(paste0("^", v), colnames(res$X), value = T)), fixed = TRUE)))
    splt <- factor(splt, labels = labs)
    df_plot <- do.call(rbind, lapply(levels(splt), function(l){
      data.frame(Value = as.vector(pred[splt == l, ]),
                 Group = l)
    }))
    meds <- tapply(df_plot$Value, df_plot$Group, median)
    postplot <- ggplot(df_plot, aes(x = Value, y = Group)) +
      geom_density_ridges(scale = 1, alpha = 0) +
      geom_segment(data = data.frame(x = meds, xend = meds, y = 1:length(meds), yend = (2:(length(meds)+1))), aes(x = x, xend = xend, y = y, yend = yend), inherit.aes = F, colour = "skyblue3", size = 1) +
      theme_bw() +
      scale_y_discrete(expand =c(0,0))+
      labs(title = p,
           x = paste0("Posterior predicted effect size"),
           y = v)
    ggsave(filename = paste0("group_plot_", v, "_", p, ".png"), plot = postplot, device = "png")
  }
}
```
